---
title: "Assingment 7"
author: "Surenther"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---
### Import CSV
```{r}
# Import CSV
c_data <- read.table(file = "student-survey.csv", header = TRUE, sep = ",")
```

### Plots
```{r warning=FALSE, message=FALSE}
# Load ggplot2
library(ggplot2)

# Plot 1: TimeReading vs TimeTV
ggplot(c_data,aes(x = TimeReading, y = TimeTV)) + geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "TimeReading vs TimeTV", x = "TimeReading", y = "TimeTV") +
  theme_minimal()

# Plot 2: TimeReading vs Happiness
ggplot(c_data,aes(x = TimeReading, y = Happiness)) + geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "green") +  
  labs(title = "TimeReading vs Happiness", x = "TimeReading", y = "Happiness") +
  theme_minimal()

# Plot 3: TimeTV vs Happiness
ggplot(c_data,aes(x = TimeTV, y = Happiness)) + geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "red") +  
  labs(title = "TimeTV vs Happiness", x = "TimeTV", y = "Happiness") +
  theme_minimal()
```
\
__Findings__ \
1. __TimeReading vs TimeTV:__ Indicates a Negative relationship, as the more time spent reading decreases the time watching TV. \
2. __TimeReading vs Happiness:__ The trend seems mixed, but there might be a weak negative relationship overall. \
3. __TimeTV vs Happiness:__ There's a Positive correlation; as time watching TV increases, happiness tends to increase. \

```{r, echo=FALSE,warning=FALSE, message=FALSE}
# Create the dataset
data <- data.frame(
  TimeReading = c(c_data$TimeReading),
  TimeTV = c(c_data$TimeTV),
  Happiness = c(c_data$Happiness)
)
# Library for visual
library(corrplot)
```

### Covariance Matrix
```{r}
#Create the correlation matrix
cov_matrix <- cov(data)
print(cov_matrix)
```
__Findings__ \
__Diagonal Values__ : These are the variances of the individual variables

* TimeReading : 3.05
* TimeTV : 174.09
* Happiness : 185.45

__Off-diagonal values__: These represent the covariances between pairs of variables \
1. __Covariance between TimeReading and TimeTV__ is -20.36, which indicates negative relationship (as reading time increases, TV time decreases) \
2. __Covariance between TimeReading and Happiness__ is -10.35, which indicates negative relationship (as reading time increases, happiness decreases) \
3. __Covariance between TimeTV and Happiness__ is 114.377, which indicated positive relationship (as TV time increase, Happiness increases as well) \

### Correlation Matrix
```{r}
# Create the correlation matrix
cor_matrix <- cor(data)
print(cor_matrix)

#Visualizing Correlation Matrix
corrplot(cor_matrix)
```
\
__Findings__ \
1. __TimeReading vs TimeTV__: The correlation between time spent reading and watching TV is -0.883, indicating a strong negative relationship. As reading time increases, TV time tends to decrease significantly. \
2. __TimeReading vs Happiness:__ The correlation between reading time and happiness is -0.43, suggesting a slight negative relationship. While there is some trend, it is not strongly negative, indicating mixed results. \
3. __TimeTV vs Happiness:__ The correlation between TV time and happiness is 0.63, reflecting a moderately positive relationship. More time spent watching TV is associated with an increase in happiness. \

### Covarience vs Correlation Matrix
Correlation Matrix is easier to interpret because it gives standardized values between -1 and 1, making it easy to understand the strength and direction of relationships between variables. Covariance Matrix provides raw co-variation, the varying scales make it harder to interpret. Thus, correlation is generally more user-friendly for understanding relationships.

### Correlation Test
```{r}
# Perform correlation test
cor_test_result <- cor.test(data$TimeReading, data$TimeTV,method="pearson")
print(cor_test_result)
```
__Findings__ \
1. __Correlation Coefficient (cor = -0.883):__ This value indicates a strong negative correlation between TimeReading and TimeTV. As the time spent on reading increases, the time spent watching TV tends to decrease significantly.\
2. __p-value = 0.0003153:__ Since the p-value is much smaller than 0.05, the correlation is statistically significant. This means there's strong evidence that the relationship between TimeReading and TimeTV is not due to random chance.\
3. __Confidence Interval (-0.969 to -0.602):__ The 95% confidence interval for the correlation is quite narrow and does not include 0, further confirming a strong negative relationship.\

Based on the correlation test, while TimeReading and TimeTV have a strong negative correlation (meaning as reading time increases, TV time decreases), this does not imply causality. Correlation alone shows a relationship but cannot confirm an effect or direction of influence between the variables.

