---
title: "Week 10 Assingment"
author: "Surenther"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

### Load ARFF file
```{r}
library(foreign)
myfile <- read.arff("ThoraricSurgery.arff")
```

### Logistic Regression
```{r}
# Convert 'Risk1Yr' to a factor for binary logistic regression
myfile$Risk1Yr <- factor(myfile$Risk1Yr, levels = c("F", "T"))

# Fit a binary logistic regression model
mymodel <- glm(Risk1Yr ~ DGN + PRE4 + PRE5 + PRE6 + PRE7 + PRE8 + PRE9 + 
               PRE10 + PRE11 + PRE14 + PRE17 + PRE19 + PRE25 + PRE30 + 
               PRE32 + AGE, data = myfile, family = binomial)
summary(mymodel)
```
__Findings__\
_From the summary, the variables with the most significant effect on survival (Risk1Yr) are:_ \ 

_PRE9T (Dyspnoea before surgery True): This predictor has a coefficient of 1.368 and a p-value of 0.00494, indicating a statistically significant positive impact on survival._\

_PRE14OC14 (clinical TNM Large): With a coefficient of 1.653 and a p-value of 0.00668, this variable also shows a strong positive influence on survival._ \

_PRE17T (Type 2 DM - diabetes mellitus True): This predictor has a coefficient of 0.9266 and a p-value of 0.03709, indicating a positive and statistically significant effect on survival._ \

_PRE30T (Smoking True): This variable has a coefficient of 1.084 and a p-value of 0.02984, suggesting it positively influences survival._\
    
_These variables showed the lowest p-values, indicating they are the most influential predictors in the model for determining one-year survival post-surgery._\

### Predict the accuracy
```{r}
# Predict probabilities
predicted_probs <- predict(mymodel, type = "response")

# Convert probabilities to binary predictions with a 0.5 threshold
predicted_classes <- ifelse(predicted_probs > 0.5, "T", "F")

# Calculate accuracy
accuracy <- mean(predicted_classes == myfile$Risk1Yr)
accuracy
```
__Findings__\
_An accuracy of 0.836 (or 83.6%) means that the logistic regression model correctly predicted the survival outcome (Risk1Yr) after surgery for approximately 83.6% of the patients in the dataset. This suggests that the model is reasonably effective at distinguishing between patients who survived one year post-surgery_