---
title: "Assingment 3"
author: "Surenther"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

### Import CSV
```{r}
# Import CSV
c_data <- read.table(file = "acs-14-1yr-s0201.csv", header = TRUE, sep = ",")
```

### Field & Data Type
```{r}
# Creating vectors
field <- c("Id","Id2","Geography","PopGroupID","POPGROUP.display.label","RacesReported","HSDegree","BachDegree")
type <- c("varchar","numeric","character","numeric","character","numeric","numeric","numeric")
intent <- c("Unique Identifier of each row","Numeric version of ID. Anonther unique Identifier","County Name","unique identifier of measures","Name of the measures (Population)","Population value","Percentage of Highschool passout","Percentage of Bachelor degree")
# Creating data frames
dt <- data.frame(Field=field,Type=type,Intent=intent)
# Library for converting into Tables
library(kableExtra)
# Converting into Table
dt %>% kbl() %>% kable_styling(bootstrap_options = c("striped", "hover"))
```

### Functions Output
```{r}
#Str output
str(c_data)
#nrow output
nrow(c_data)
#ncol output
ncol(c_data)
```

### Histogram
```{r}
#Load ggplot2
library(ggplot2)
#Create basic ggplot
basic <- ggplot(c_data,aes(x=HSDegree))
#Add histogram
hist <- basic + geom_histogram(color=4,fill="white", bins = 50) 
#Add title and axis lable
title <- hist + ggtitle("Distribusion of High School Degree") + labs(y = "Count", x = "HS Degree") 
#Add theme
graph <- title + theme_minimal()
#Display 
graph
```

### Histogram Questions
1. Is the data distribution unimodal? \
No, this distribution is not unimodal. It exhibits multiple peaks within the 85% to 95% range.

2. Is it approximately symmetrical? \
No, the distribution is not approximately symmetrical. The graph shows a right skew, where most of the data is concentrated between 85% and 95%.

3. Is it approximately bell-shaped? \
No, the distribution is not approximately bell-shaped. A bell-shaped distribution is symmetric and has a single peak, resembling a normal distribution. In this case, the distribution has multiple peaks.

4. Is it approximately normal? \
No, the distribution is not approximately normal. A normal distribution is symmetric, bell-shaped, and has a single peak in the center, with data tapering off equally on both sides.

5. If not normal, is the distribution skewed? If so, in which direction? \
Yes, the distribution is skewed to the left. This means most of the data is on the higher end (around 85% to 95%), while the lower end has fewer values, creating a longer tail on the left side.

6. Include a normal curve to the Histogram that you plotted.
```{r}
#Load ggplot2
library(ggplot2)
#Create basic ggplot
basic <- ggplot(c_data,aes(x=HSDegree))
#Add histogram with density
hist <- basic + geom_histogram(aes(y = after_stat(density)),color=4,fill="white", bins = 50)
# Add normal distribution line
line <- hist + stat_function(fun = dnorm, args = list(mean = mean(c_data$HSDegree), sd = sd(c_data$HSDegree)), color = "red", linewidth = 1)
#Add title and axis label
title <- line + ggtitle("Distribusion of High School Degree with normal curve") + labs(y = "Density", x = "HS Degree") 
#Add theme
graph <- title + theme_minimal()
#Display 
graph
```

7. Explain whether a normal distribution can accurately be used as a model for this data. \
Based on the histogram with the normal curve overlay, it appears that the data is not perfectly normal, which means a normal distribution may not be the best model for this data.

### Probability Plot
```{r}
# Load ggplot2
library(ggplot2)
# Create basic ggplot
basic <- ggplot(c_data,aes(sample=HSDegree))
# Q-Q plot
qq <- basic + stat_qq()
# Reference line
line <- qq + stat_qq_line(color = "red")
#Add title and axis label
title <- line + ggtitle("Probability Plot") + labs(y = "HS Degree", x = "")
#Add theme
graph <- title + theme_minimal()
#Display 
graph
```

### Probability Plot Questions
1. Based on what you see in this probability plot, is the distribution approximately normal? Explain how you know. \
Based on the probability plot, the distribution does not appear to be perfectly normal. In a normal distribution, the data points would closely follow the straight red line. Here, the points deviate from the line at the lower and upper ends

2. If not normal, is the distribution skewed? If so, in which direction? Explain how you know. \
The distribution appears to be right-skewed.

### stat.desc() 
```{r}
#Load pastecs
library(pastecs)
stat.desc(c_data[c('RacesReported','HSDegree','BachDegree')], basic=TRUE, desc=TRUE, norm=FALSE, p=0.95)
```

### skew, kurtosis, and z-scores
Skew: Skewness measures how much the data is asymmetrical. A positive skew means more data points are concentrated on the left with a long tail on the right. A negative skew means the opposite.

Kurtosis: Kurtosis tells us about the "peakedness" of the data. High kurtosis means more data is in the tails (extreme values), while low kurtosis means fewer extreme values.

Z-scores: Z-scores show how far each data point is from the mean in terms of standard deviations. A z-score of 2 means the point is 2 standard deviations away from the mean.

Sample size impact: With a larger sample size, results for skewness, kurtosis, and z-scores become more reliable. Small samples can lead to random variation, making the data appear more skewed or have higher kurtosis than it truly is.