---
title: "Assignment 5"
author: "Surenther"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

### Load XLSX file
```{r}
#Load XLSX
library(readxl)
mydata <- read_excel("week-6-housing.xlsx", 1)
```

## Dplyr Functions

### Mutate - Create Year column
```{r}
#Extract the Year column from the Date and Create a new column
library(magrittr) #Library for Pipes
library(dplyr,warn.conflicts = FALSE) 
mydata2 <- mydata
mydata2 %<>%
  select('Sale Date','Sale Price','zip5','square_feet_total_living','bedrooms') %>%
  mutate('Sale Year' = format(mydata$'Sale Date','%Y'))
print(mydata2,width=Inf,n=5)
```

### Summarize - Avg Sale Price, Avg Sqft, Avg Bedroom
```{r}
#Calculate Avg Saleprice, SQFT and Bedroom
mydata2 %>%
  summarize_at(
   vars(AvgSalePrice="Sale Price",AvgSqft="square_feet_total_living",  
        AvgBedroom="bedrooms"),mean, na.rm = TRUE
   )
```

### Group by - Yearsold, Zip
```{r}
#Group by Year,Zip & Calculate Avg Saleprice, SQFT and Bedroom
mydata2 %>%
  group_by( 
    Year = mydata2$'Sale Year',Zip = mydata2$'zip5'
    ) %>% 
  summarize_at(
    vars(AvgSalePrice="Sale Price",AvgSqft="square_feet_total_living",  
         AvgBedroom="bedrooms"),mean, na.rm = TRUE
    )
```

### Select - Sale Date, Sale Price, Zip & Square Footage
```{r}
#Select Sale data, Sale Price, Zip and Square feet
mydata %>%
  select('Sale Date','Sale Price','zip5','square_feet_total_living')
```

### Filter - based on Sale Year and Price and Sqft
```{r}
#Find home with more than 3500sqft with less than 650k sold at 2016
mydata2 %>%
  print(filter(mydata2$'Sale Year' == "2016" & mydata2$'Sale Price' <= 650000  
               & mydata2$'square_feet_total_living' > 3500),width=Inf,n=5)
```

### Arrange - by Avg Saleprice in desc
```{r}
#Group by Year,Zip & Calculate Avg Saleprice, SQFT and Bedroom, Display by Avg Saleprice 
mydata2 %>%
  group_by( 
    Year = mydata2$'Sale Year',Zip = mydata2$'zip5'
    ) %>% 
  summarize_at(
    vars(AvgSalePrice="Sale Price",AvgSqft="square_feet_total_living",  
         AvgBedroom="bedrooms"),mean, na.rm = TRUE
    ) %>%
  arrange(desc(AvgSalePrice))
```

## purrr Functions
### keep - Only Numeric Columns 
```{r}
library(purrr,warn.conflicts = FALSE)
numeric_data <- keep(mydata, is.numeric)
numeric_data
```
Findings: It removed non numeric columns like Sale Date,addr_full,ctyname etc

### discard - Remove Colunmn with NA
```{r}
cleaned_data <- discard(mydata, ~ any(is.na(.)))
cleaned_data
```
Findings: It removed columns like sale_warning,ctyname which is having missing value

### compact - Remove NULL values from the data
```{r}
clean_without_null <- compact(mydata)
clean_without_null
```

## Cbind - Column Bind
```{r}
# Extract Home details from the data
home_details <- 
  mydata %>%
  select(year_built,square_feet_total_living,bedrooms)
# Extract Sale details from the data
sale_detail <-
  mydata %>%
  select('Sale Date','Sale Price',sale_reason)
#Cbind
c_merge <- cbind(home_details,sale_detail)
head(c_merge,n=5)
```

## Rbind - Row Bind
```{r}
#2006 Sale details
sale_2006 <-
  mydata2 %>%
  filter(mydata2$'Sale Year' == "2006")
#2016 Sale details
sale_2016 <-
  mydata2 %>%
  filter(mydata2$'Sale Year' == "2016")
#Rbind
r_merge <- rbind(sale_2006,sale_2016)
head(r_merge,n=5)
```

## Split- Extract house no & Street name from address
```{r}
library(stringr)
#Split the full address based on first space
mydata[c('Home_No', 'Street Name')]  <- str_split_fixed(string=mydata$addr_full,  
                                                        pattern=" ",2)
#display the data
mydata %>% select(addr_full,Home_No,'Street Name')
```

## concatenate - address with proper format 
```{r}
mydata %>%
  mutate('modified_addr' = sprintf("%s,%s %s-%s",mydata$Home_No,mydata$'Street Name',  
                                   mydata$postalctyn,mydata$zip5)) %>%
  select(addr_full,modified_addr)
```